VariantSpark: Population Scale Clustering of Genotype Information

### 2015-09-17 doc/publications/journal2014/paper.tex
Figure 3 uses PDF with legend
added citation

{\sc Apache Spark}~\cite{Zaharia2011} is a more recent compute engine, which overcomes many of Hadoop's limitations. 
One of the main benefits is that it allows programs to cache data in memory; potentially eliminating, or at least reducing, the bottleneck of disk IO. 
When utilising caching, Apache claim {\sc Spark} to be 100x faster than Hadoop. 
Although {\sc Spark} allows MapReduce-like programs, it does not require programs to exactly model the MapReduce paradigm, which in turn allows more flexibility when designing programs. 
Recognising the capability, the Big Data Genomics (BDG) group recently demonstrated the strength of {\sc Spark} in a genomic clustering application using {\sc ADAM}, a set of formats and APIs as well as processing stage implementations for genomic data~\cite{Massie2013}. 
While the speedup over traditional methods was impressive (50 fold speedup), being limited by constraints within this general genomics framework hampered performance.

>>>>

{\sc Apache Spark}~\cite{Zaharia2011} is a more recent compute engine, which overcomes many of Hadoop's limitations. 
One of the main benefits is that it allows programs to cache data in memory; potentially eliminating, or at least reducing, the bottleneck of disk IO. 
When utilising caching, Apache claim {\sc Spark} to be 100x faster than Hadoop. 
Although {\sc Spark} allows MapReduce-like programs, it does not require programs to exactly model the MapReduce paradigm, which in turn allows more flexibility when designing programs. 
Recognising the capability, Wiewi{\'o}rka {\it et al.}~\cite{Wiewiorka2014} developed SparkSeq for high-throughput sequence data analysis and the Big Data Genomics (BDG) group recently demonstrated the strength of {\sc Spark} in a genomic clustering application using {\sc ADAM}, a set of formats and APIs as well as processing stage implementations for genomic data~\cite{Massie2013}. 
While the speedup of {\sc ADAM} over traditional methods was impressive (50 fold speedup), being limited by constraints within this general genomics framework hampered performance.
